### 2025-01-23 paper/main.tex
Updates from Overleaf

\paragraph{Abstract:}
It is known that any permutation-invariant function of data points $r_i$
can be written in the form $\rho(\sum_i\phi(r_i))$,
where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
This form---known in the machine-learning literature as Deep Sets---is a map--reduce form, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
The area of a triangle is a permutation-invariant function of the locations $r_i$ of the three corners $1\leq i\leq 3$.
We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
It is the square-root of a permutation-invariant quartic polynomial.
No insights of any kind were gained by this work.

@@

@@

@@

>>>>

\paragraph{Abstract:}
It is known that any permutation-invariant function of data points $r_i$
can be written in the form $\rho(\sum_i\phi(r_i))$,
where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
This form---known in the machine-learning literature as Deep Sets---is a map--reduce form, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
The area of a triangle is a permutation-invariant function of the locations $r_i$ of the three corners $1\leq i\leq 3$.
We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
It is the square-root of a permutation-invariant quartic polynomial involving the integer 368.
No insights of any kind were gained from this work.

@@

HAINJE: Given data of form $(x, y)$, invariant factors are $F_{m,n} = \sum_i x_i^m y_i^n$, and the invariant monomials are any products of these factors $\prod_{i} F_{m_i,n_i}$. Do we want/need to be more specific than that?

@@

HAINJE: Here it is:
\begin{gather}
    \phi(r_i) = (x_i, y_i, x_iy_i, x_i^2, y_i^2, x_i y_i^2, x_i^2 y_i, x_i^2 y_i^2) \\
    \begin{aligned}
    \rho(\textstyle\sum_{i=1}^{3}\phi(r_i)\equiv \vec{u}) = &
    - \tfrac{54}{368} u_8
    - \tfrac{258}{368} u_3^2
    + \tfrac{285}{368} u_4 u_5
    + \tfrac{36}{368} u_2 u_7
    + \tfrac{36}{368} u_1 u_6 \\ &
    - \tfrac{101}{368} u_2^2 u_4
    + \tfrac{148}{368} u_1 u_2 u_3
    - \tfrac{101}{368} u_1^2 u_5
    + \tfrac{9}{368} u_1^2 u_2^2.
    \end{aligned}
\end{gather}

@@

\newpage
By inspection, we have also found a simpler form of the same formula, which is given by
\begin{align}
    4 \, A^2 = & \
        2 \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i ) \, (\sum_{i=1}^{3} x_i \, y_i )
        + 3 \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 ) \nonumber \\ &
        - \, (\sum_{i=1}^{3} x_i )^2 \, (\sum_{i=1}^{3} y_i^2 )
        - \, (\sum_{i=1}^{3} y_i )^2 \, (\sum_{i=1}^{3} x_i^2 ) \nonumber \\ &
        - 3 \, (\sum_{i=1}^{3} x_i \, y_i )^2.
\end{align}
In explicit Deep Sets notation, this can be written
\begin{gather}
    \phi(r_i) = (
        x_i,
        y_i,
        x_i y_i,
        x_i^2,
        y_i^2
    ), \\
    \rho(\textstyle\sum_{i=1}^{3} \phi(r_i) \equiv \vec{u})
    = \tfrac{2}{4} u_1 u_2 u_3
    + \tfrac{3}{4} u_4 u_5
    - \tfrac{1}{4} u_1^2 u_5
    - \tfrac{1}{4} u_2^2 u_4
    - \tfrac{3}{4} u_3^2.
\end{gather}

The two are related by
\begin{equation}
    4 u_1 u_6 + 4 u_2 u_7 - 6 u_8 = 
    -u_1^2 u_2^2 + 4 u_1 u_2 u_3 - 2 u_3^2 + u_2^2 u_4 + u_1^2 u_5 - u_4 u_5 ;
\end{equation}
namely, plugging this into \eqref{eq:result} causes it to reduce immediately into this simpler form.

This form was not found by the symbolic regression of \secref{sec:method} because the regression makes use of \verb+lstsq+ with more parameters than data points. In this regime, the solution that minimizes the $\mathcal{L}_2$ norm of the weights (the coefficients of the invariant monomials) is returned. The $\mathcal{L}_2$ norm of the weights of our initial formula is $2127 / 1472 \approx 1.445$, whereas the $\mathcal{L}_2$ norm of the weights of this simpler formula is $3/2 = 1.5$.

HAINJE: Would lasso regression return my formula? Is there a yet simpler formula?
